Promise and peril: Agnostic identification methods\\ for detecting differential item functioning

### 2020-03-14 paper/paper_1_intro.Rmd
printed second verison!

<!-- need to specify which multidimensional model im talking about -->

Inspired by @camilli1992conceptual, we think differential item functioning (DIF) as bias against a group of students that manifests when an item response model with too few ability dimensions is imposed. From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is a property of the student, not the item. For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability. In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.

Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, @kopf2015framework simulate students as belonging to either a reference or focal group. They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values from a previous study. They set item easinesses for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} + 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits. They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is $$P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)$$ where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.

We find it informative to describe simulation conditions in terms of multidimensional abilities as opposed to item parameters that vary across groups. For example, to translate the @kopf2015framework simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously ${b_j}^{\text{ref}}$ for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for the reference group and ${\eta_i}^{\text{ref}} = -1$ for the focal group. We now need to use a 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional 2PL model, the probability student $i$ responding correctly to item $j$ is
$$
P(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
$$
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability for students in the same group.

>>>>

Inspired by @camilli1992conceptual, we think of differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed. From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item. For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability. In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.

Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, @kopf2015framework simulate students as belonging to either a reference or focal group. They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values from a previous study. They set item easinesses for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits. They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is $$P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)$$ where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.

We find it informative to describe simulation conditions in terms of multidimensional abilities as opposed to item parameters that vary across groups. For example, to translate the @kopf2015framework simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously ${b_j}^{\text{ref}}$ for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for all students in the reference group and ${\eta_i}^{\text{ref}} = -1$ for all students in the focal group. We now need to use a compensatory 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional compensatory 2PL model, the probability student $i$ responding correctly to item $j$ is
$$
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
$$
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in nuisance ability across students in the same group.

### 2020-03-14 paper/paper_2a_anchor_items.Rmd
printed second verison!

Psychometricians have long been in search of the perfect DIF detection method. In this section, we summarize existing DIF methods and propose extensions of those methods. We take as given that the IRT-based likelihood ratio test is the best way to test for DIF, which has been well demonstrated [@meade2012solving]. As a result, we avoid methods like the Mantel-Haenszel procedure [@holland1986differential], which muddies the waters by moving away from the IRT framework, and has been shown to perform no better — and in some cases worse — than IRT-based methods [@swaminathan1990detecting].

The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions — about either relative group ability or which items contain bias — can be made. For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection. And, in search of a coherent framework, we sometimes edit names of existing methods. We recognize that others have done the same [e.g. @kopf2015framework] and that we risk contributing to the proliferation of names.

We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test. Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ and ${\eta_i}^{\text{ref}} = -1$. The slope on target ability is set to $a_{1j} = 1$ for all items, and the slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise. From the DIF-as-item-property view, there is no nuisance ability. Instead, ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for the first five items and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5$ for the last three items.

In general, we denote the mean ability of the reference and focal group as $\mu^\text{ref}$ and $\mu^\text{foc}$, respectively. We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$. The fundamental challenge for DIF methods is to disentangle estimating $\hat\mu^\text{foc}$ from estimating $\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$. The most common approach is to identify a group of anchor items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$.

## Anchor items

@@

@meade2012solving compared the most commonly used methods and unequivocally recommended the all-others-as-anchors (AOAA). AOAA tests each item for DIF one at a time using all of the other items as anchors. For example, when testing the first item for DIF, all of the other items are used as anchors. This is done using a likelihood ratio test that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups [@thissen1993detection]. Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on. The items for which the flexible model outperforms the baseline model (typically based on a chi-squared test) are identified as having DIF and the rest of the items become anchor items.

@edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF. How can we be sure its those 10 items and not the other 11 items with DIF? We can't be. Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible. More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group and others are biased against the reference group. Undesirably, most applications of DIF methods and many simulation studies do not make explicit the assumptions of the DIF method [@strobl2018anchor]. In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.

@@

One way to attempt to counter artificial DIF is with purification through iterative anchor selection. For example, @drasgow1987study started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively — with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model — until no more items tested positively. @kopf2015framework named this technique Iterative-backward-AOAA with "backward" (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free. We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS). Appending all-significant indicates that anchor selection is done iteratively with all items that show statistical significance for DIF being removed from the anchor set.

AOAA-AS might seem like an improvement but it doesn't solve a fundamental problem of AOAA: What do we do when all items test positive for DIF? With a sufficient sample size and at least one item with DIF, this will necessarily be the case. In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF and there are no anchor items. @kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly. @woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF. It's unclear if one should proceed if those four items display DIF themselves.

@@

We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed. AOAA-OAT is inspired by @hagquist2017recent who, in general, assert that "items showing DIF initially
should not be resolved simultaneously but sequentially" [p. 6]. Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the $\chi^2$ test statistic, is removed from the anchor set. The process continues iteratively until no new items display DIF. AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF. As a result, AOAA-OAT is less likely to arrive at the wrong conclusion because of artifical DIF.

@@

Table \ref{table:themethods} summarizes the three all-others-as-anchors methods. It's useful to remember that AOAA is not an iterative procedure, while the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method.

@@

\begin{table}[H]
\caption{EDIT THIS}
\centering
\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
\toprule

@@

[ADD IN SIMULATION RESULTS]

@@

@bechger2015statistical proposed choosing anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider relative item parameters. For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$. The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in (easiness) between items" [p. 323].

The general idea of identifying clusters of items is intriguing. However, their approach is needlessly complicated and they did not describe a process for moving from $\Delta \mathbf{R}$ to a set of anchor items. @pohl2017cluster extended their work by proposing one such process. $\Delta \mathbf{R}$ is skew-symmetric and of rank 1, which means that all information is contained in a single row or column. Accordingly, they use k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC. They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster. Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.

We propose a new cluster-based approach, which we call "equal means clustering" (EMC). Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$ of differences in item easiness, $\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}$, where the model is identified by setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, thus the name "equal means clustering"). This results in all differences in performance — either from group ability differences or DIF — manifesting in the item easiness difference parameter $\tilde{d_j}$. The tilde is used to denote parameters estimated with $\mu^\text{foc}$ arbitrarily set to $0$.

Instead of choosing k with BIC, we use the gap statistic method recommended by @hastie2001estimating. The largest cluster is chosen as the anchor cluster. If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected. This process makes the assumption that the largest cluster (but necessarily the majority) of items do not contain DIF.

OUR SIMULATION RESULTS

```{r}
out_3_biased_items$cluster_status[[1]] %>% 
    select(item, a_ref_easy, b_foc_easy, difference_in_easy, cluster) %>% 
    mutate(anchor = ifelse(cluster == 1, TRUE, FALSE)) %>% 
    kable()
```

### Analyst chooses

DIF methods are generally designed to automatically detect DIF without any human judgement. On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.

We propose a technique, the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst. Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc}$ to $0$. Again, the result is that all differences in performance manifest in $\tilde{\mathbf{d}}$. To get a sense of variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical]. Then, multiple imputations (MI) [@yang2012characterizing] are used to estimate the distribution of $\tilde{d_j}$ for every item. These are the distributions displayed in the EM-MILG. The method is inspired in part by @pohl2017cluster who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and @talbot2013taking who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.

Figure \ref{fig:emmilg} shows the EM-MILG for our simulated data. It cannot be stated strongly enough that the EM-MILG contains all possible information about the difference in group performance. The challenge, then, is to select the anchor items. The analyst might assume that — because there are five items where the reference group outperforms the focal group by 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and can be used as anchor items.

```{r emmilg, fig.cap = 'The equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.', out.width="70%", warning = FALSE, message = FALSE}
out_3_biased_items$intuitive_mod[[1]] %>%
    mod_intuitive_to_draws_df() %>%
    draws_df_to_logit_plot() +
    labs(
        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
    )

@@

After choosing the anchor items, the model is refit. The new model is identified by setting $d_j = 0$ for the anchor items, instead of by setting $\mu^\text{foc}$ to $0$. The same process of using multiple imputations to estimate the distribution of $d_j$ can be used with the new model. Because the equal means assumption is not made, we refer to the resulting graph as a multiple imputations logit graph (MILG). In our simulation, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.

```{r milg, fig.cap = 'The multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed to 0.', out.width="70%", warning = FALSE, message = FALSE}
# CAREFUL - this code assumes that the analyst would choose the same anchor items as AOAA-OAT found

@@

One of our key concerns with typical DIF methods is that they can lull the analyst into a false sense of security. Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items. The EM-MILG combats this concern by presenting all information clearly to the analyst. In the previous example, the analyst may be weary of their results having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased. Even when other DIF methods are used, the analyst can use the EM-MILG as a first step in order to give them a sense of their item response data. And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG.

>>>>

Psychometricians have long been in search of the perfect DIF detection method. It has become well known that the IRT-based likelihood ratio test (LRT) is the best way to test for DIF, which has been well demonstrated [@meade2012solving]. As a result, we avoid methods like the Mantel-Haenszel procedure [@holland1986differential], which muddies the waters by moving away from the IRT framework, and has been shown to perform no better — and in some cases worse — than IRT-based methods [@swaminathan1990detecting].

The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions — about either relative group ability or which items contain bias — can be made. That is, we're interested in assumptionless identification methods (hereafter referred to as "AI methods") so that the LRTs can be used to test items for DIF. For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection. And, in search of a coherent framework, we sometimes edit names of existing methods. We recognize that others have done the same [e.g. @kopf2015framework], and that we risk contributing to the proliferation of names.

In this section, we summarize existing AI methods and propose extensions of those methods. We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test. Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ and ${\eta_i}^{\text{ref}} = -1$. The slope on target ability is set to $a_{1j} = 1$ for all items. The slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise. From the DIF-as-item-property view, there is no nuisance ability. Instead, ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for the first five items and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5$ for the last three items.

In general, we denote the mean ability of the reference and focal group as $\mu^\text{ref}$ and $\mu^\text{foc}$, respectively. We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$. The fundamental challenge is to use an AI method to identify the model, which allows for the disentangling $\hat\mu^\text{foc}$ from estimating $\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$. The most common AI method is to select a group of anchor items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$, and the remaining items can be tested for DIF using an LRT.

Computing is done in R [@rcore], model fitting in the R package mirt [@chalmers2012mirt], and data wrangling/visualization in the set of R packages known as the tidyverse [@tidy]. Code is available at TO DO.

## Anchor items {#anchoritems}

@@

@meade2012solving compared the most commonly used AI methods and unequivocally recommended the all-others-as-anchors (AOAA) method. AOAA tests each item for DIF one at a time using all of the other items as anchors. For example, when testing the first item for DIF, all of the other items are used as anchors. This is done using an LRT that compares the baseline model, where all item parameters are fixed across groups, to the flexible model, where the parameters of the tested item are freed across groups [@thissen1993detection]. Then, when testing the second item for DIF, once again all of the other items (including the first item) are used as anchors, and so on. The items for which the flexible model outperforms the baseline model (typically based on a $\chi^2$ test) are identified as having DIF, and the rest of the items become anchor items. Note that AOAA is implemented in the mirt R package, and is called by passing scheme = "drop" to the DIF function (drop refers to dropping a single constraint when moving from the constrained to the flexible model).

@edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF. How can they be sure that its those 10 items and not the other 11 items with DIF? They cannot be. Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible. More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group. Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method [@strobl2018anchor]. In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.

@@

One way to attempt to counter artificial DIF is with purification through iterative anchor selection. For example, @drasgow1987study started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively — with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model — until no more items tested positively. @kopf2015framework named this technique Iterative-backward-AOAA with "backward" (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free. We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS). Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set. Note that AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.

AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF? With a sufficient sample size and at least one item with DIF, this will necessarily be the case. In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items. @kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly. @woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF. It's unclear if one should proceed if those four items display DIF.

@@

We propose an extension of these methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), which, to our knowledge (and surprise), has not previously been explicitly proposed. AOAA-OAT is inspired by @hagquist2017recent, who, in general, assert that "items showing DIF initially should not be resolved simultaneously but sequentially" [p. 6]. Like AOAA-AS, AOAA-OAT starts with AOAA, but only the single item exhibiting the most DIF, based on the $\chi^2$ test statistic, is removed from the anchor set. The process continues iteratively until no new items display DIF. AOAA-OAT and AOAA-AS are similar in that they are both iterative; the difference is that AOAA-OAT takes the more conservative approach of removing only one item in each iteration as opposed to all items that test positive for DIF. As a result, we believe that AOAA-OAT is less likely to be "fooled" by artifical DIF. Note that AOAA-OAT is not currently implemented in the R package mirt.

@@

Table \ref{table:themethods} summarizes the three all-others-as-anchors methods. It's useful to remember that AOAA is not an iterative procedure. The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being our new, more conservative method. In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified. On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.

@@

\begin{table}[H]
\caption{Summary of the three all-others-as-anchors assumptionless identification methods}
\centering
\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
\toprule

@@

@@

@bechger2015statistical proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider only relative item parameters. For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$. The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in (easiness) between items" [p. 323].

The general idea of identifying clusters of items is intriguing. However, their approach is needlessly complicated, and they did not describe a process for moving from $\Delta \mathbf{R}$ to an anchor cluster. @pohl2017cluster extended their work by proposing one such process. $\Delta \mathbf{R}$ is skew-symmetric and of rank 1, which means that all information is contained in a single row or column. Accordingly, they use k-means clustering on just the first column of $\Delta \mathbf{R}$ where the number of clusters, k, is chosen by minimizing BIC. They suggest using a combination of cluster size, cluster homogeneity, and cluster parameter precision to choose which of the clusters is the anchor cluster. Unfortunately, in their simulation study, they find that BIC identifies only a single cluster, so they end up using all of the items as anchors.

We propose a new cluster-based approach, which we call "equal means clustering" (EMC). Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$ of differences in item easiness, $\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}$, where the model is identified by setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, thus the name "equal means clustering"). As a result, all differences in performance — either from group ability differences or DIF — manifesting in the item easiness difference parameter $\tilde{d_j}$. The tilde is used to denote parameters estimated with $\mu^\text{foc}$ arbitrarily set to $0$.

Instead of choosing k with BIC, we use the gap statistic method recommended by @hastie2001estimating. The largest cluster is chosen as the anchor cluster. If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected. This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF. In our one-run simulation, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.

### The equal means, multiple imputation logit graph

AI methods are generally designed to automatically detect DIF without any human judgement. On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.

We propose a new method, the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst. Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc}$ to $0$. Again, the result is that all differences in performance manifest in $\tilde{\mathbf{d}}$. To measure the variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical]. Then, multiple imputations (MI) [@yang2012characterizing] are used to estimate the distribution of $\tilde{d_j}$ for each item. These are the distributions displayed in a EM-MILG. The method is inspired in part by @pohl2017cluster who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and @talbot2013taking who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.

Figure \ref{fig:emmilg} shows the EM-MILG for our simulated data. We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance. The challenge, then, is to select the anchor items. The analyst might assume that — because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and can be used as anchor items.

```{r emmilg, fig.cap = 'A equal means multiple imputations logit graph (EM-MILG) shows the distribution of how many logits the reference group outperforms the focal group by on each item.', out.width="70%", warning = FALSE, message = FALSE}
out_3_biased_items$intuitive_mod[[1]] %>%
    mod_intuitive_to_draws_df() %>%
    draws_df_to_logit_plot() +
    labs(
        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{ref} - \\tilde{b_j}^{foc}$")
    )

@@

After choosing the anchor items, the model is refit. The new model is identified by setting $d_j = 0$ for the anchor items, instead of by setting $\mu^\text{foc}$ to $0$. The same process of using multiple imputations to estimate the distribution of $d_j$ can be used with the new model. Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG). As expected, in our one-run simulation, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.

```{r milg, fig.cap = 'A multiple imputations logit graph (MILG) shows the distribution of DIF against the focal group. Anchor items are fixed by setting $d_j = 0$.', out.width="70%", warning = FALSE, message = FALSE}
# CAREFUL - this code assumes that the analyst would choose the same anchor items as AOAA-OAT found

@@

One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security. Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items. A EM-MILG combats this concern by presenting all information clearly to the analyst. In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased. Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data. And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG. In particular, we could visualize DIF in the non-anchor items after using AOAA or CLUSTER to identify the model.
